---
title: "rs_project"
output: html_document
date: "2023-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=F message=F}
library(dplyr)
library(readxl)
library(sf) # replaces "sp", used for vector
library(terra) # replaces "raster"
library(R.utils)
library(ggplot2)
library(tidyr)
library(cowplot)
library(randomForest)
library(yardstick)
#library(RStoolbox)

# define projection for project
NAD83_UTM13 <- "EPSG:6342"
```

# Read in field data and W. RMNP boundary
```{r}
# read in ground-truth data
raw <- read_excel("field_data/DATA.xlsx", sheet = "point info")
metrics <- read.csv("field_data/mpb_severity_metrics_perovich.csv")
df <- inner_join(raw, metrics, by=c("Site"="PLOT.ID"))

# convert to vector points
df_sf <- st_as_sf(df, coords=c("X", "Y"), crs=NAD83_UTM13)

# read in the shape file representing the region of interest for clipping the data
extent_wrmnp <- read_sf("rmnp_boundary/qgis_layers/rmnp_west_side/rmnp_west_side.shp")

# read in the shape file representing the area of prediction (pico forest)
romo_veg_map <- read_sf("romo_veg_map/veg_map_utm/veg_map_utm.shp") 
```

```{r}
# try another y-var

# % dead ba (all spp, all ages) <- rodman
trees1 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4) %>%
  mutate(dead = if_else(STATUS %in% c(3,5), 1, 0)) %>%
  group_by(`PLOT ID`, dead) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarize(dead_ba_percent = ba[dead==1]/sum(ba)*100) %>%
  ungroup()

# % dead ba (all spp, adults)
trees2 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4 & DBH > 5) %>%
  mutate(dead = if_else(STATUS %in% c(3,5), 1, 0)) %>%
  group_by(`PLOT ID`, dead) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarise(dead_ad_ba_percent = ba[dead==1]/sum(ba)*100) %>%
  ungroup()

#3 % dead or MPB ba (all spp, any age)
trees3 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4) %>%
  mutate(dead_or_mpb = if_else(STATUS %in% c(3,5) | MPB == 1, 1, 0)) %>%
  group_by(`PLOT ID`, dead_or_mpb) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarize(dead_or_mpb_ba_percent = ba[dead_or_mpb==1]/sum(ba)*100) %>%
  ungroup()

#4 % dead or MPB ba (all spp, adults)
trees4 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4 & DBH > 5) %>%
  mutate(dead_or_mpb = if_else(STATUS %in% c(3,5) | MPB == 1, 1, 0)) %>%
  group_by(`PLOT ID`, dead_or_mpb) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarize(dead_or_mpb_ad_ba_percent = ba[dead_or_mpb==1]/sum(ba)*100) %>%
  ungroup()

trees5 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4) %>%
  mutate(dead = if_else(STATUS %in% c(2,3,5), 1, 0)) %>%
  group_by(`PLOT ID`, dead) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarize(dead_or_unhealthy_ba_percent = ba[dead==1]/sum(ba)*100) %>%
  ungroup()

# % dead ba (all spp, adults)
trees6 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4 & DBH > 5) %>%
  mutate(dead = if_else(STATUS %in% c(2,3,5), 1, 0)) %>%
  group_by(`PLOT ID`, dead) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarise(dead_or_unhealthy_ad_ba_percent = ba[dead==1]/sum(ba)*100) %>%
  ungroup()

#3 % dead or MPB ba (all spp, any age)
trees7 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4) %>%
  mutate(dead_or_mpb = if_else(STATUS %in% c(2,3,5) | MPB == 1, 1, 0)) %>%
  group_by(`PLOT ID`, dead_or_mpb) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarize(dead_or_unhealthy_or_mpb_ba_percent = ba[dead_or_mpb==1]/sum(ba)*100) %>%
  ungroup()

#4 % dead or MPB ba (all spp, adults)
trees8 <- 
  read_excel("field_data/DATA.xlsx", sheet = "trees") %>%
  filter(STATUS != 4 & DBH > 5) %>%
  mutate(dead_or_mpb = if_else(STATUS %in% c(2,3,5) | MPB == 1, 1, 0)) %>%
  group_by(`PLOT ID`, dead_or_mpb) %>%
  dplyr::summarize(ba = sum(DBH^2 * 0.005454)) %>%
  ungroup() %>% 
  group_by(`PLOT ID`) %>% 
  dplyr::summarize(dead_or_unhealthy_or_mpb_ad_ba_percent = ba[dead_or_mpb==1]/sum(ba)*100) %>%
  ungroup()

df <- 
  df %>%
  inner_join(trees1, by=c("Site"="PLOT ID")) %>%
  inner_join(trees2, by=c("Site"="PLOT ID")) %>%
    inner_join(trees3, by=c("Site"="PLOT ID")) %>%
    inner_join(trees4, by=c("Site"="PLOT ID")) %>%
    inner_join(trees5, by=c("Site"="PLOT ID")) %>%
    inner_join(trees6, by=c("Site"="PLOT ID")) %>%
    inner_join(trees7, by=c("Site"="PLOT ID")) %>%
    inner_join(trees8, by=c("Site"="PLOT ID"))


```

# function to geoprocess landsat images
```{r}
process_landsat <- function(filepath, filename_prefix){
    
  # get filenames
  filenames <- paste0(filepath, filename_prefix, "_SR_B", c(1:5,7), ".TIF")
  qa_filename <- paste0(filepath, filename_prefix, "_QA_PIXEL.TIF")
  
  # read in spectral bands and QA layer
  L <- terra::rast(filenames)
  names(L) <- c('blue', 'green', 'red', 'NIR', 'SWIR1', 'SWIR2') 
  L_qa <- terra::rast(qa_filename)
  names(L_qa) <- "qa"
  
  # set buffer, to speed up reprojection in next step
  buff <- buffer(vect(extent_wrmnp), 10000) 
  
  # geoprocess SR bands
  L <- 
    crop(L, buff) %>%         # clip image to speed up reprojection
    project(NAD83_UTM13) %>%  # reproject
    mask(extent_wrmnp)        # crop to the ROI
  
  # more geoprocessing...  
  L <- 
    (L * 0.0000275 - 0.2) %>%
    clamp(0, 1) 

  # geoprocess qa raster
  L_qa <-
    crop(L_qa, buff) %>%      # clip image to speed up reprojection
    project(NAD83_UTM13) %>%  # reproject
    mask(extent_wrmnp)        # crop to the ROI
  
  # QA_PIXEL is generated by the CFMask algorithm. Get values to be masked:
  cloud_dilations <- bitwAnd(L_qa[], 2^1) > 0  # bit1=cloud dilation
  clouds <- bitwAnd(L_qa[], 2^3) > 0           # bit3=clouds
  cloud_shadows <- bitwAnd(L_qa[], 2^4) > 0    # bit4=clouds
  
  # assign non-pico pixels to NA in landsat image
  L_qa[cloud_dilations] <- NA
  L_qa[clouds] <- NA
  L_qa[cloud_shadows] <- NA 
  
  # mask out clouds in imagery bands
  L_clear <- mask(L, L_qa)
  
  # return L_clear
  L_clear
}  
```

# Read and geoprocess landsat images
```{r, eval=F}

######################
# 2009 L5 image
post <- process_landsat(filepath="landsat/L5_20090821/", 
                        filename_prefix="LT05_L2SP_034032_20090821_20200827_02_T1")

# 2010 L5 image
post <- process_landsat(filepath="landsat/L5_20100925/", 
                        filename_prefix="LT05_L2SP_034032_20100925_20200823_02_T1")

# 2009 L5 image
post <- process_landsat(filepath="landsat/L5_20090821/", 
                        filename_prefix="LT05_L2SP_034032_20090821_20200827_02_T1")

# 2000 L5 image
pre <- process_landsat(filepath="landsat/L5_20000913/", 
                        filename_prefix="LT05_L2SP_034032_20000913_20200906_02_T1")

# 1998 L5 image
pre <- process_landsat(filepath="landsat/L5_19980908/", 
                        filename_prefix="LT05_L2SP_034032_19980908_20200908_02_T1")
```
```{r}
######
# 100% cloud-free - performs best
post_date <- "9/17/2010"
post <- process_landsat(filepath="landsat/L7_20100917/",  filename_prefix="LE07_L2SP_034032_20100917_20200910_02_T1")

# 99% cloud-free - performs best
pre_date <- "8/2/1999"
pre <- process_landsat(filepath="landsat/L7_19990802/",   filename_prefix="LE07_L2SP_034032_19990802_20200918_02_T1")

# 100% clear - SWIR bands are really correlated, but probably deflated bc of springtime moisture. 
#post_date <- "6/18/2012"
#post <- process_landsat(filepath="landsat/L7_20120618/",  filename_prefix="LE07_L2SP_034032_20120618_20200908_02_T1")

# 99% cloud-free
#post_date <- "9/9/2007"
#post <- process_landsat(filepath="landsat/L7_20070909/",  filename_prefix="LE07_L2SP_034032_20070909_20200913_02_T1")

# 93% cloud-free
#pre_date <- "7/19/2000"
#pre <- process_landsat(filepath="landsat/L7_20000719/", filename_prefix="LE07_L2SP_034032_20000719_20200918_02_T1")

# 87% cloud-free
#post_date <- "7/15/2010"
#post <- process_landsat(filepath="landsat/L7_20100715/", filename_prefix="LE07_L2SP_034032_20100715_20200911_02_T1")

# 95% cloud-free 
#pre_date <- "8/18/1999"
#pre <- process_landsat(filepath="landsat/L7_19990818/",filename_prefix="LE07_L2SP_034032_19990818_20200918_02_T1")

#pre_date <- "6/23/2002"
#pre <- process_landsat(filepath="landsat/L7_20020623/",filename_prefix="LE07_L2SP_034032_20020623_20200916_02_T1")

#post_date <- "7/15/2010"
#post <- process_landsat(filepath="landsat/L7_20100715/", filename_prefix="LE07_L2SP_034032_20100715_20200911_02_T1")

```

# Get PICO forest in W. RMNP. This the area we'll predict over
```{r}
# geoprocess ROMO veg map to get a raster of pico forests in our ROI
pico_rast <- 
  romo_veg_map %>%
  filter(COMMON_MAP %in% c(
    'Lodgepole Pine - Low Elevation < 9500 ft', 
    'Lodgepole Pine - High Elevation > 9500 ft', 
    'Lodgepole Pine - Rock')) %>%  
  rasterize(y=post$NIR) %>%   # shouldn't matter which landsat image I use
  mask(extent_wrmnp)          # crop to the ROI

# clip landsat to pico stands only
post_pico <-  mask(post, pico_rast)  
pre_pico <- mask(pre, pico_rast)

# plot for funsies (true color)
plotRGB(pre_pico, b=1, g=2, r=3, stretch="hist") 
plotRGB(post_pico, b=1, g=2, r=3, stretch="hist") 

# false color
plotRGB(pre_pico, b=2, g=4, r=5, stretch="hist") 
plotRGB(post_pico, b=2, g=4, r=5, stretch="hist") 

# total pico area
area_pico <- length(pico_rast[!is.na(pico_rast)])*30^2/1000^2  

# % cloudless (I can only do this on one band, so I arbitrarily chose NIR)
length(pre_pico$NIR[!is.na(pre_pico$NIR)])*30^2/1000^2/area_pico*100  # pre
length(post_pico$NIR[!is.na(post_pico$NIR)])*30^2/1000^2/area_pico*100  # post
```


# Extract SR values to field points. These are the points we'll train on
```{r}

# plot field points over prediction area
plotRGB(pre_pico, b=1, g=2, r=3, stretch="hist") 
points(df_sf, col="red", cex=1)


# functionalize...
sr_to_points <- function(rast){

    # convert field points to spatvector
    df_vect <- vect(df_sf)

    # extract spectral reflectance for each field point
    terra::extract(rast, df_vect, method="bilinear") %>% # use 4 nearest raster cells
    bind_cols(df) %>%
    select(dead_or_mpb_ba_percent, blue:SWIR2) %>%
    filter_all(all_vars(!is.na(.))) 
}

post_points <- sr_to_points(post) 
pre_points <- sr_to_points(pre) 


```

```{r}
# quantify cloud cover of field points
points_clear <- function(points_df){
  
  # total field points 
  all_points <- length(vect(df_sf))  # 219 plots
  
  # cloudless field points (arbitrarily chose NIR)
  cloudless_points <- 
    points_df %>%
    filter(NIR != "NaN") %>%
    count() %>%
    pull()
  
  # % of points that are cloudless
  cloudless_points/all_points*100 
}

points_clear(post_points)
points_clear(pre_points)
```
# determine classification cutoff
```{r}
kmeans_clusters <- kmeans(post_points$dead_or_mpb_ba_percent, centers = 3)

# classify low-med-high MPB severity
post_points$mpb_severity <- kmeans_clusters$cluster

#post_points <- 
 # post_points %>%
 # mutate(mpb_severity = case_when(mpb_severity == 1 ~ "low",
 #                                 mpb_severity == 2 ~ "med",
 #                                 mpb_severity == 3 ~ "high")) %>%
 # mutate(mpb_severity = case_when(mpb_ba_percent < 60 ~ "low",
#                                  mpb_ba_percent >= 60 & mpb_ba_percent <= 80 ~ "med",
#                                  mpb_ba_percent > 80 ~ "high")) %>%
#  mutate(mpb_severity = factor(mpb_severity, levels=c("low","med","high")))
  
post_points <- 
  post_points %>%
  mutate(mpb_severity = case_when(dead_or_mpb_ba_percent <= 59.4 ~ "low",
                                  dead_or_mpb_ba_percent > 59.4 & dead_or_mpb_ba_percent <= 79.9 ~ "med",
                                  dead_or_mpb_ba_percent > 79.9 ~ "high")) %>%
  mutate(mpb_severity = factor(mpb_severity, levels=c("low","med","high")))


# apply to pre_points (this'll drop the points that are masked out of post_points)
num_cat_xwalk <- 
  post_points %>% 
  select(dead_or_mpb_ba_percent, mpb_severity) %>%
  distinct() 

pre_points <- inner_join(pre_points, num_cat_xwalk, by = 'dead_or_mpb_ba_percent')


```


# Plot spectral signatures
```{r}
# plot spectral signatures
# get band values for field points (df_sr = output of sr_to_points)
plot_spectral_sig <- function(df_sr, title){
  
 # set band order
 band_order <- c("blue", "green", "red", "NIR", "SWIR1", "SWIR2")
 
   df_sr %>%
   pivot_longer(cols=blue:SWIR2, names_to="band") %>%
   mutate(band = factor(band, levels=band_order))  %>%
   ggplot(aes(x=band, y=value, fill=mpb_severity)) +
   geom_boxplot() +
    ylim(0,.2) +
    labs(y="Reflectance",
        title=title) 
}

p_post <- plot_spectral_sig(post_points, title=paste("Post-Outbreak", post_date))
p_pre <- plot_spectral_sig(pre_points, title=paste("Pre-Outbreak", pre_date))

plot_grid(p_pre, p_post, ncol=1)
```


```{r}
# do a boxplot of the difference, fill by mpb_severity

plot_spectral_sig_diff <- function(df_sr_pre, df_sr_post, title){
  
 # set band order
 band_order <- c("blue", "green", "red", "NIR", "SWIR1", "SWIR2")
 
   df_sr_pre %>%
   pivot_longer(cols=blue:SWIR2, names_to="band") %>%
   mutate(band = factor(band, levels=band_order))  %>%
   ggplot(aes(x=band, y=value, fill=mpb_severity)) +
   geom_boxplot() +
    ylim(0,.2) +
    labs(y="Reflectance",
        title=title) 
}
```

interesting, the pre-outbreak green is lower...maybe there's recovery in the post-outbreak image? Also, why are trees already separated in SWIR in the pre-outbreak. Drought?  

Remember:
* SWIR captures leaf moisture (lower = moister) Recent precip can affect this
* Foliage reflects NIR (to avoid overheating), though conifers reflect less
* Chlorophyll reflects green 

from 1999 to 2000 to 2002, we see increasingly more RBG (??), less NIR (less foliage), and more SWIR (droughtier)
from 2010 to 2012, we see more reflectance in all bands. More foliage, droughtier...? we do see a better separation between outbreak severities in 2010 than 2012

- NDMI between 0 and -0.20: low canopy cover or low moisture content
- NDMI values vary throughout the growing season because the plants’ reflectance is slightly different for every phenological stage (not sure how true this is for conifers). 
- There also exists an interesting correlation between NDMI and NDVI. Water stress indicated by the NDMI values can be confirmed by a significantly lower than average NDVI.


# calculate pre and post NDMI in W. RMNP (not just picos)
```{r}
#####################
# calculate dNDMI
pre_ndmi <- (pre$NIR - pre$SWIR1)/(pre$NIR + pre$SWIR1)
post_ndmi <- (post$NIR - post$SWIR1)/(post$NIR + post$SWIR1)
dNDMI <- post_ndmi - pre_ndmi
names(dNDMI) <- "dNDMI"
  
# plot dNDMI
plot(dNDMI, main="dNDMI")

#####################
# calculate dNDWI
pre_ndwi <- (pre$green - pre$SWIR1)/(pre$green + pre$SWIR1)
post_ndwi <- (post$green - post$SWIR1)/(post$green + post$SWIR1)
dNDWI <- post_ndwi - pre_ndwi
names(dNDWI) <- "dNDWI"
  
# plot dNDWI
plot(dNDWI, main="dNDWI")

#####################
# calculate dNDVI
pre_ndvi <- (pre$NIR - pre$red)/(pre$NIR + pre$red)
post_ndvi <- (post$NIR - post$red)/(post$NIR + post$red)
dNDVI <- post_ndvi - pre_ndvi
names(dNDVI) <- "dNDVI"
  
# plot dNDVI
plot(dNDVI, main="dNDVI")

#####################
# calculate simple index
pre_vi <- pre$SWIR1/pre$NIR
post_vi <- post$SWIR1/post$NIR
dVI <- post_vi - pre_vi
names(dVI) <- "dVI"
  
# plot dVI
plot(dVI, main="dVI")

#####################
# calculate dVARI
pre_vari <- (pre$green-pre$red)/(pre$green + pre$red - pre$blue )
post_vari <- (post$green-post$red)/(post$green + post$red - post$blue )
dVARI <- post_vari - pre_vari
names(dVARI) <- "dVARI"

#####################
# calculate dTCBRI
pre_tcbri <-  (0.2043*pre$blue) + (0.4158*pre$green ) +(0.5524*pre$red) + (0.5741*pre$NIR) + (0.3124*pre$SWIR1) + (0.2303*pre$SWIR2)
post_tcbri <-  (0.2043*post$blue) + (0.4158*post$green ) +(0.5524*post$red) + (0.5741*post$NIR) + (0.3124*post$SWIR1) + (0.2303*post$SWIR2)
dTCBRI<- post_tcbri - pre_tcbri
names(dTCBRI) <- "dTCBRI"


#####################
# calculate dTCGRE
pre_tcgre <-  (-0.1603*pre$blue) + (-0.2819*pre$green ) +(0.4934*pre$red) + (0.7940*pre$NIR) + (-0.0002*pre$SWIR1) + (-0.1446*pre$SWIR2)
post_tcgre <-  (-0.1603*post$blue) + (-0.2819*post$green ) +(0.4934*post$red) + (0.7940*post$NIR) + (-0.0002*post$SWIR1) + (-0.1446*post$SWIR2)
dTCGRE <- post_tcgre - pre_tcgre
names(dTCGRE) <- "dTCGRE"

#####################
# calculate dTCWET
pre_tcwet <-  (0.0315*pre$blue) + (0.2021*pre$green ) +(0.3102*pre$red) + (0.1594*pre$NIR) + (-0.6806*pre$SWIR1) + (-0.6109*pre$SWIR2)
post_tcwet <-  (0.0315*post$blue) + (0.2021*post$green ) +(0.3102*post$red) + (0.1594*post$NIR) + (-0.6806*post$SWIR1) + (-0.6109*post$SWIR2)
dTCWET <- post_tcwet - pre_tcwet
names(dTCWET) <- "dTCWET"



dRED <- post$red - pre$red
names(dRED) <- "dRED"
dGREEN <- post$green - pre$green
names(dGREEN) <- "dGREEN"
dBLUE <- post$blue - pre$blue
names(dBLUE) <- "dBLUE"
dNIR <- post$NIR - pre$NIR
names(dNIR) <- "dNIR"
dSWIR1 <- post$SWIR1 - pre$SWIR1
names(dSWIR1) <- "dSWIR1"
dSWIR2 <- post$SWIR2 - pre$SWIR2
names(dSWIR2) <- "dSWIR2"

#####################

# convert field points to spatvector
df_vect <- vect(df_sf)

# bilinear: use 4 nearest raster cells
dVIs <-
  terra::extract(dNDMI, df_vect, method="bilinear") %>% 
  inner_join(terra::extract(dNDWI, df_vect, method="bilinear")) %>% 
  inner_join(terra::extract(dNDVI, df_vect, method="bilinear")) %>% 
  inner_join(terra::extract(dVI, df_vect, method="bilinear")) %>% 
  inner_join(terra::extract(dVARI, df_vect, method="bilinear")) %>%
  inner_join(terra::extract(dTCBRI, df_vect, method="bilinear")) %>%
  inner_join(terra::extract(dTCGRE, df_vect, method="bilinear")) %>%
  inner_join(terra::extract(dTCWET, df_vect, method="bilinear"))
  #inner_join(terra::extract(dRED, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(dGREEN, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(dBLUE, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(dNIR, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(dSWIR1, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(dSWIR2, df_vect, method="bilinear")) %>%
  #inner_join(terra::extract(post$blue, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(post$green, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(post$red, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(post$NIR, df_vect, method="bilinear")) %>% 
  #inner_join(terra::extract(post$SWIR1, df_vect, method="bilinear")) %>%
  #inner_join(terra::extract(post$SWIR2, df_vect, method="bilinear"))  %>%
 


```

Compare other dVIs: VARI, NDVI, NDWI. Plot against dNDMI to check for correlation
```{r}
# hmm they're all pretty correlated
p1 <- ggplot(dVIs)+geom_point(aes(x=dVI, y=dNDWI))
p2 <- ggplot(dVIs)+geom_point(aes(x=dVI, y=dNDVI))
p3 <- ggplot(dVIs)+geom_point(aes(x=dNDVI, y=dNDWI))
plot_grid(p1, p2, p3)

ggplot(dVIs)+geom_point(aes(x=dNDVI, y=dNDMI))


```

Classify with RF (xgboost is better) 
```{r}

# extract spectral reflectance for each field point
dVI_points <- 
  bind_cols(dVIs, df) %>%
  inner_join(num_cat_xwalk, by = 'dead_or_mpb_ba_percent') %>%
  select(mpb_severity, dNDMI:dTCWET, Elevation) %>% 
  mutate(mpb_severity = factor(mpb_severity, levels=c("high","med","low"))) %>%
  filter_all(all_vars(.!="NaN")) %>% 
  group_by(mpb_severity) %>%
  mutate(mpb_severity = factor(mpb_severity, levels=c("low", "med", "high"))) %>% 
  filter(!is.na(Elevation)) %>%
  data.frame()


ggplot(dVI_points) + geom_boxplot(aes(x=dNDMI, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dNDWI, y=mpb_severity))  # 
ggplot(dVI_points) + geom_boxplot(aes(x=dNDVI, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dVI, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dRED, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dGREEN, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dBLUE, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dNIR, y=mpb_severity))   #
ggplot(dVI_points) + geom_boxplot(aes(x=dSWIR1, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dSWIR2, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=blue, y=mpb_severity))  
ggplot(dVI_points) + geom_boxplot(aes(x=green, y=mpb_severity))  
ggplot(dVI_points) + geom_boxplot(aes(x=red, y=mpb_severity))  
ggplot(dVI_points) + geom_boxplot(aes(x=NIR, y=mpb_severity))  
ggplot(dVI_points) + geom_boxplot(aes(x=SWIR1, y=mpb_severity))  
ggplot(dVI_points) + geom_boxplot(aes(x=SWIR2, y=mpb_severity))  #
ggplot(dVI_points) + geom_boxplot(aes(x=Elevation, y=mpb_severity)) #
ggplot(dVI_points) + geom_boxplot(aes(x=dTCWET, y=mpb_severity)) ##
ggplot(dVI_points) + geom_boxplot(aes(x=dTCBRI, y=mpb_severity)) 
ggplot(dVI_points) + geom_boxplot(aes(x=dTCGRE, y=mpb_severity)) 




```

```{r}
###### Numeric Regression ###### 

# extract spectral reflectance for each field point - numeric style
dVI_points_num <- 
  bind_cols(dVIs, df) %>%
  select(dead_or_mpb_ba_percent, dNDMI:dvari, Elevation) %>% 
  filter_all(all_vars(.!="NaN")) %>% 
  filter(!is.na(Elevation)) %>%
  data.frame()

# plot y vs VI
ggplot(dVI_points_num) +
  geom_point(aes(x=dead_or_mpb_ba_percent, y=dNDMI)) +
  labs(x="% BA dead")

# correlation matrix
dVI_points_mat <- as.matrix(dVI_points_num)
library(Hmisc)
rcorr <- rcorr(dVI_points_mat) 
y_rcorr <- rcorr$r["dead_or_mpb_ba_percent", ] %>% round(2)
print(y_rcorr)

# split training/testing data
train <- sample_frac(dVI_points_num, 0.7, replace = FALSE)
test <- setdiff(dVI_points_num, train) %>% data.frame()

set.seed(10)

# build RF model
mod <- randomForest(dead_or_mpb_ba_percent ~ .,
                    data = train,
                    importance=TRUE, 
                    ntree=200,
                    mtry=5)

mod

# regression
lm(dead_or_mpb_ba_percent ~ ., data=dVI_points_num) %>% summary()

```

```{r}
# split training/testing data
train <- sample_frac(dVI_points, 0.7, replace = FALSE)
test <- setdiff(dVI_points, train) %>% data.frame()

set.seed(300)

# build RF model
mod <- randomForest(mpb_severity ~ .,
                    data = train,
                    importance=TRUE, 
                    ntree=200,
                    mtry=5)

mod

# CHECK ERROR CONVERGENCE
plot(mod)

# PLOT mean decrease in accuracy VARIABLE IMPORTANCE
# expresses how much accuracy (GINI coeff) the model losses by excluding each var
varImpPlot(mod, type=1)
```


```{r}
# focused model
mod <- randomForest(mpb_severity ~ dNDMI,
                    data = train,
                    importance=TRUE, 
                    ntree=100)
mod

# testing accuracy
pred <- predict(mod, test)
forCF <- data.frame(ref=test$mpb_severity, pred=pred)
conf_mat(forCF, ref, pred)
summary(conf_mat(forCF, ref, pred))

```

```{r}
# map predictions
library(tmap)

pre_pico_ndmi <- (pre_pico$NIR - pre_pico$SWIR1)/(pre_pico$NIR + pre_pico$SWIR1)
post_pico_ndmi <- (post_pico$NIR - post_pico$SWIR1)/(post_pico$NIR + post_pico$SWIR1)
dNDMI_pico <- post_pico_ndmi - pre_pico_ndmi
names(dNDMI_pico) <- "dNDMI"
plot(dNDMI_pico) 


rastPred <- terra::predict(object=dNDMI_pico, model=mod)

tm_shape(rastPred) +
  tm_raster(style= "cat",title="Land Cover")

f <- freq(rastPred)
f$area <- f$count * 30^2/1000^2 
f$percent = round(100 * f$area / sum(f$area), 1)
f
```


```{r}
p_nirswir_pre <- ggplot(pre_points) + geom_point(aes(x=SWIR1, y=NIR, color=mpb_severity)) + labs(title="Pre")
p_nirswir_post <- ggplot(post_points) + geom_point(aes(x=SWIR1, y=NIR, color=mpb_severity)) + labs(title="Post")

plot_grid(p_nirswir_post, p_nirswir_pre)

```
